---
title: "Bayesian Distributions"
author: "William Duquette and Soham Changani"
date: "12/18/2021"
output: html_document
---
### Making of Bayesian Distributions

```{r setup, include=FALSE}
#Import Neccesary Libraries
require(car)
require(lattice)
library(HH)
require(mosaic)
require(ggplot2)
require(tidyverse)
require(EnvStats)
require(MASS)
library(MuMIn)
require(gridExtra)
require(Hmisc)
library(expss)
library(ResourceSelection)
require(effsize)
library(asbio)
library(Stat2Data)
require(dplyr)
require(fBasics)
require(nortest)
library(lmtest)
library(nlme)
library(sandwich)
require(agricolae)
require(DescTools)
require(BSDA)
```


The function below creates a very simple Bernoulli likelihood function. A value for theta and data is passed in and it creates a simple likelihood function. Theta is a probability, meaning that it must range from 0 to 1. Whereas the data is a vector of data.
```{r}
bernoulli_likelihood <- function(theta, data) {
  
  n <- length(data)
  z <- sum(data)
  return(theta^z * (1 - theta)^(n - sum(data)))
  
}
```


Example: Prior based on a relatively small sample size (Not enough to shift posterior)
```{r}
n <- 12# Number of observations (Bernoulli)
z <- 9 # Number of "successes" (Bernoulli)
a <- 2 # Number of "successes" (Beta)
b <- 10 # Number of "failures" (Beta)

trial_data <- c(rep(0, times = n - z), rep(1, times = z)) #creates vector containing N-z zeros and z zeros

#Creation of all three distributions: Beta prior, Bernoulli Likelihood, and Beta posterior 
d <-tibble(theta = seq(from = 0, to = 1, length.out = 100)) %>% 
  mutate(`Prior (beta)` = dbeta(theta, shape1 = a, shape2 = b),
         `Likelihood (Bernoulli)` = bernoulli_likelihood(theta = theta, data  = trial_data),
         `Posterior (beta)`= dbeta(theta, shape1 = z+a, shape2 = n-z+b))
```

```{r}
#Graph distributions
d %>% 
  gather(key, value, -theta) %>% 
  mutate(key = factor(key, levels = c("Prior (beta)", "Likelihood (Bernoulli)", "Posterior (beta)"))) %>% 
  
  #Plot
  ggplot(aes(x = theta)) +geom_ribbon(aes(ymin = 0, ymax = value), fill = "blue") +labs(x = expression(theta)) +    
  facet_wrap(~key, scales = "free_y", ncol = 1) + theme(panel.grid = element_blank())
```



Example: Prior based on a relatively large sample size (Shifts posterior)
```{r}
n <- 12 # Number of observations (Bernoulli)
z <- 9 # Number of "successes" (Bernoulli)
a <- 21 # Number of "successes" (Beta)
b <- 99 # Number of "failures" (Beta)

trial_data <- c(rep(0, times = n - z), rep(1, times = z)) #creates vector containing N-z zeros and z zeros

#Creation of all three distributions: Beta prior, Bernoulli Likelihood, and Beta posterior 
d <-tibble(theta = seq(from = 0, to = 1, length.out = 100)) %>% 
  mutate(`Prior (beta)` = dbeta(theta, shape1 = a, shape2 = b),
         `Likelihood (Bernoulli)` = bernoulli_likelihood(theta = theta, data  = trial_data),
         `Posterior (beta)`= dbeta(theta, shape1 = z+a, shape2 = n-z+b))
```


```{r}
#Graph distributions
d %>% 
  gather(key, value, -theta) %>% 
  mutate(key = factor(key, levels = c("Prior (beta)", "Likelihood (Bernoulli)", "Posterior (beta)"))) %>% 
  
  #Plot
  ggplot(aes(x = theta)) +geom_ribbon(aes(ymin = 0, ymax = value), fill = "blue") +labs(x = expression(theta)) +    
  facet_wrap(~key, scales = "free_y", ncol = 1) + theme(panel.grid = element_blank())
```

